---
title: "Metcalf_Data_Replication_Assignment"
author: "Cat Metcalf"
date: "2023-10-26"
output: html_document
---

# Data Replication Assignment

This replication assignment is based on an article by Munro et al. that discusses the development of agriculture through human management in the Southern Levant.

## Background

## Supplementary NISP data from Table 1, p. 3
(May not have to include)

First we must load in the data sets that provide the supplementary data:
```{r}
library(curl)
library(tidyverse)
library(ggplot2)

f <- curl("https://raw.githubusercontent.com/cattuccino/Metcalf-data-replication-assignment/main/AN588%20DataRep%20SuppData%202.csv")
S3 <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(S3)

```

## Replication of Figure 2 with a Polynomial Regression Line and Spearman's Rho Calculation, p. 5
Plots the relative abundance indices and calculates a Spearman's Rho with associated p-value (Descriptive analysis and 2 Inferential Analyses).

Uses supplementary data: Figure 2a uses Table S3, 2b uses Table S4, 2c uses S5, and 2d uses S5.

First we have to make a plot of the data: (do for each plot)
```{r}
library(dplyr)
```

```{r}
gazella_med <- 
  S3 |>
  select("SITE", "PERIOD", "X.Gazella") |>
  filter(PERIOD == "KEB" | PERIOD == "GKEB" | 
           PERIOD == "ENAT" | PERIOD == "LNAT" | 
           PERIOD == "PPNA" | PERIOD == "EPPNB" | 
           PERIOD == "MPPNB") |>
  arrange("PERIOD") 

```

```{r}
gazella_med |>
  ggplot(data = gazella_med, 
  mapping = aes(x = as.factor(PERIOD), y = X.Gazella)) + 
  geom_pointrange(stat = "summary",
                  fun = mean,
                  fun.min = function(z){quantile(z, 0.25)},
                  fun.max = function(z){quantile(z, 0.75)}) +
  labs(y = "Proportion Gazelle") +
  scale_x_discrete(limits = c("KEB", "GKEB", "ENAT", "LNAT", "PPNA", "EPPNB", "MPPNB")) +
  theme_bw()
```



## Replication of Abundance Indices in Chronological Order Figure 2, p. 6
(May end up removing but may do just to learn for myself)

Edit